---
title: "lab_02"
author: "katie benzan"
date: "2/4/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

**Task** Run the codeblock below to load the Tidyverse and Lubridate. To run the codeblock, hit the little green play button in the upper right corner of the gray area (the codeblock) below.

```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse and lubridate
library(tidyverse)
library(lubridate)
```

## Load Data

Let's keep working with West Virginia loans contained in the PPP loans data.  

**Task** Run the codeblock below to load the data.

```{r}
west_virginia_ppp <- read_rds("data/lab_02.rds")
```

## Examine Data

## Answer questions

**Q1** Create a new column that divides the loan amounts into the following options - under_10k, 10k_50k, 50k_100k, 100k_500k, 500k_1m, 1m_plus - and calculates how many loans are in each group. How would you describe the distribution of loans in a sentence?
Hint: You will need to use `case_when` to make this work.
**A1** The majority of loans were between $10,000 and $50,000, whereas the fewest loans were $1,000,000 or more. The group ranging from $10,000 to $50,000 was almost 33% larger than the next largest group, which is $10,000. 

```{r}
# Write the code to produce your answer here
west_va_categories <- west_virginia_ppp %>%
mutate(
  loan_amounts_category = case_when(
    amount < 10000 ~ 'under_10k',
    amount >= 10000 & amount < 50000 ~ '10k_50k',
    amount >= 50000 & amount < 100000 ~ '50k_100k',
    amount >= 100000 & amount < 500000 ~ '100k_500k',
    amount >= 500000 & amount < 1000000 ~ '500k_1m',
    amount >= 1000000 ~ '1m_plus'
  )
)

west_va_categories %>%
  group_by(loan_amounts_category) %>%
  summarise(count = n())
```

**Q2** Some of the loans appear not to have been paid out - there is a column representing the amount of money that hasn't been disbursed to the applicant. Which servicing lenders have the most loans with undisbursed amounts greater than 0?
**A2** Davis Trust Company had the most loans with undisbursed amounts greater than 0. 

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>%
  filter(undisbursed_amount > 0) %>%
  group_by(servicing_lender_name) %>% 
  summarise(
    count = n()
    ) %>%
arrange(desc(count))
```

**Q3** What are some potential reasons why a loan might not be paid out? What questions does that circumstance raise for you?
**A3** A loan might not be paid out because their income may not have increased or stabilized as predicted. It also may seem that a bank may be more willing to lend out money to riskier companies than other banks. By the data above, it appears as though Davis Trust Company was willing to lend money to companies unable to pay back their loans. 
- something went wrong on the borrower's end OR the lender decided not to follow through b/c they don't trust


**Q4** Which month in 2021 had the most loans approved?
**A4** March 2021

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>%
  filter(date_approved > '2020-12-31') %>% 
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>% 
  summarise(count = n()) %>% 
  arrange(desc(count))
  
```

**Q5** How many loans went to West Virginia applicants that listed the maximum number of jobs retained (500)? Which one had the earliest approved date?
**A5** 10 loans went to West Virginia applicants with 500 jobs retained. J. AUSTIN HEATING & COOLING, INC. had the earliest approved date - April 8, 2020.

```{r}
# Write the code to produce your answer here
west_virginia_ppp %>%
  filter(jobs_retained == 500) %>% 
  arrange(date_approved)
```
