---
title: "Class Reference"
author: "Katie Benzan"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this my personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

to set working directory: more --> set as working directory
desired working directory must be open

to run code --> the green play button in the top right corner of the gray code box

install.packages() - to install packages
install.packages('tidyverse') - to install all of tidyverse at once
install.packages('rmarkdown') - to install R markdown

readr - library that reads different types of data (csv)
library(readr)
dplyr - analyzes that data
library(dplyr)
'package:stats' - filter, lag
'package:base' - intersect, setdiff, setequal, union
shortcut - library(tidyverse)

to import data: ppp_maryland_loans <- read_rds("ppp_maryland.rds")

%>% - "and then do this"
data %>% function - take data set and then do this specific action to it

n() - counts the number of rows or records in each group
arrange() - arranges data in order (default - ascending)
  - how you want to see the data
arrange(desc()) - descending order
sum() - add up
sum(amount) - add up values to a new column 'amount'
mean() - find the mean
median() - find the median
min() - lowest value
max() - highest value

head() - the first six rows in dataset 
What can I expect from this dataset?

To summarize data: 
```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    
  )
```

Then to organize summarized data:
```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    
  ) %>%
  arrange(desc(count_loans))
```

# - comment something (won't run code)

lubridate library - makes it easier to work with date-times in R
mutate - 
  mutate(new_column = value)
  mutate(percent_payroll = payroll_proceed/amount)
  mutate(name = "Katie Benzan") - change an existing column's name
filering  
dates 
str_to_upper(name) - make all uppercase
select(name) - show me

to order by descending:
```{r}
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = (payroll_proceed/amount)*100
)  %>% arrange(desc(percent_payroll))
```

order ascending percent_payroll for the largest amount:
```{r}
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = (payroll_proceed/amount)*100
)  %>% arrange(percent_payroll, desc(amount))
```

if_else - binary answer
  i.e. in Maryland or out of Maryland?
```{r}
# if/else example
maryland_with_in_out <- maryland_ppp %>%
  mutate(
    in_out = if_else(
        servicing_lender_state == 'MD', "IN", "OUT"
      )
  )
```
== - test of equivalence

case_when