---
title: "Class Reference"
author: "Katie Benzan"
output:
  html_document:
    theme: cerulean
    highlight: pygments
    toc: true
    toc_float:
      collapsed: true
      smooth_scroll: false
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

Consider this my personal guide to the commands and functions you will learn. In general, when you come across an R command or function that you want to remember, put it in here along with a description of what it does and when you'd use it.

to set working directory: more --> set as working directory
desired working directory must be open

to run code --> the green play button in the top right corner of the gray code box

install.packages() - to install packages
install.packages('tidyverse') - to install all of tidyverse at once
install.packages('rmarkdown') - to install R markdown

readr - library that reads different types of data (csv)
library(readr)
dplyr - analyzes that data
library(dplyr)
'package:stats' - filter, lag
'package:base' - intersect, setdiff, setequal, union
shortcut - library(tidyverse)

to import data: ppp_maryland_loans <- read_rds("ppp_maryland.rds")

%>% - "and then do this"
data %>% function - take data set and then do this specific action to it

n() - counts the number of rows or records in each group
arrange() - arranges data in order (default - ascending)
  - how you want to see the data
arrange(desc()) - descending order
sum() - add up
sum(amount) - add up values to a new column 'amount'
mean() - find the mean
median() - find the median
min() - lowest value
max() - highest value

head() - the first six rows in dataset 
What can I expect from this dataset?

To summarize data: 
```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    
  )
```

Then to organize summarized data:
```{r}
ppp_maryland_loans %>%
  group_by(project_county_name) %>%
  summarise(
    
  ) %>%
  arrange(desc(count_loans))
```

# - comment something (won't run code)

lubridate library - makes it easier to work with date-times in R
mutate - 
  mutate(new_column = value)
  mutate(percent_payroll = payroll_proceed/amount)
  mutate(name = "Katie Benzan") - change an existing column's name
filtering  
dates 
str_to_upper(name) - make all uppercase
select(name) - show me

to order by descending:
```{r}
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = (payroll_proceed/amount)*100
)  %>% arrange(desc(percent_payroll))
```

order ascending percent_payroll for the largest amount:
```{r}
maryland_ppp %>%
  select(loan_number, amount, payroll_proceed) %>%
  mutate(
  percent_payroll = (payroll_proceed/amount)*100
)  %>% arrange(percent_payroll, desc(amount))
```

if_else - binary answer
  i.e. in Maryland or out of Maryland?
```{r}
# if/else example
maryland_with_in_out <- maryland_ppp %>%
  mutate(
    in_out = if_else(
        servicing_lender_state == 'MD', "IN", "OUT"
      )
  )
```
== - test of equivalence

case_when
```{r}
# case_when example
maryland_jobs_categories <- maryland_ppp %>%
mutate(
  jobs_retained_category = case_when(
    jobs_retained < 10 ~ 'under_10',
    jobs_retained >= 10 & jobs_retained < 25 ~ '10_to_25',
    jobs_retained >= 25 & jobs_retained < 50 ~ '25_to_50',
    jobs_retained >= 50 & jobs_retained < 100 ~ '50_to_100',
    jobs_retained >= 100 & jobs_retained < 250 ~ '100_to_250',
    jobs_retained >= 250 & jobs_retained < 500 ~ '250_to_500',
    jobs_retained == 500 ~ '500'
  )
)
```

filters
```{r}
prince_georges <- maryland_ppp %>% filter(project_county_name == "PRINCE GEORGES")
head(prince_georges)
```
AND
```{r}
and_prince_georges <- maryland_ppp %>% 
  filter(project_county_name == "PRINCE GEORGES" & business_type == "Non-Profit Organization" & amount > 150000)

nrow(and_prince_georges)
```
OR
```{r}
or_prince_georges <- maryland_ppp %>% 
  filter(project_county_name == "PRINCE GEORGES" | business_type == "Non-Profit Organization" | amount > 150000)

nrow(or_prince_georges)
```

read.csv

to create a new column called month
```{r}
maryland_ppp %>%
  mutate(Month = floor_date(date_approved, "month")) %>%
  group_by(Month) %>%
  summarise(total = n()) %>%
  arrange(desc(total))
```

guess_max - set the dataset type by the first 10 rows (guess_max=10)
  certain amount of confidene that 10 rows is enough to determine the type
  lower number, more risk tolerant

```{r}
ppp_maryland_loans <- read_csv("ppp_loans_md.csv.zip", guess_max=10)
  # error message
```
  VS.
```{r}
ppp_maryland_loans <- read_csv("ppp_loans_md.csv.zip", guess_max=195865)
  # no error message
```

```{r}

problems(ppp_maryland_loans)
 # mismatch between what was expected and the actual value
```

```{r}
ppp_maryland_loans <- ppp_maryland_loans %>%
  mutate(id = as.character(id))
# changing the field from a number to a character value
# overriding the previous same dataset --> naming the same dataset with the same name (replacing it in its entirety)
# as.date, as.numeric, etc.
```

filter(is.na(forgiveness_date)) --> check to see if value is NA ... 

```{r}
ppp_maryland_loans %>%
  mutate(month_year_approved = floor_date(date_approved, "month")) %>%
  group_by(month_year_approved) %>%
   summarise(
    count=n()
  ) %>%
  arrange(month_year_approved)

# to create a new column called "month_year_approved" that extracts the month and year from the "date_approved" column  
# Group by the new "month_year_approved" column 
# count  
# sort by "month_year_approved"
```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names()

# display the cleaned dataset
cleaned_arnold_md_loans

# standardizes the names of columns across the board
# makes sure the column names are easily readable by R
```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id)

# renaming a column

# display the cleaned dataset
cleaned_arnold_md_loans


```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount))

# changing column from character to numeric

# display the cleaned dataset
cleaned_arnold_md_loans

```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct()

# to eliminate any duplicated rows

# display the cleaned dataset
cleaned_arnold_md_loans

```

```{r}
# cleaning function
cleaned_arnold_md_loans <- arnold_md_loans %>%
  clean_names() %>%
  rename(id = x1_id) %>%
  mutate(amount = as.numeric(amount)) %>%
  distinct() %>%
  mutate(zip = str_sub(zip, start=1L, end=5L))

# to standardize the zip column
# all to 5 digits

# display the cleaned dataset
cleaned_arnold_md_loans
