---
title: "lab_08"
author: "katie benzan"
date: "3/17/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## You will need

* A Census API key

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and tidycensus.

**Task** load these two packages

```{r}
# Turn off scientific notation
options(scipen=999)
library(tidyverse)
library(tidycensus)
```

## Setup Census API

You'll need your Census API Key:

**Task** set your API Key

```{r echo=FALSE}
census_api_key("102c0c72d8f8eb880608f9dacb53fdf9100f8ba0", overwrite = TRUE, install = TRUE)
```

## Load and view ACS variables

You'll also need to load the list of American Community Survey variables from the 2019 5-year estimates so you can identify the codes you need:

**Task** load the variables from the 2019 5-year ACS and save them to a dataframe

```{r}
# Write code here to load the ACS variables
acs_2019 <- load_variables(2019, "acs5", cache = TRUE)
View(acs_2019)
```

## Load PPP data

You'll also need to load an RDS file of PPP loan applications that were removed from the original data by lenders in July.

**Task** read the RDS file and save it to a dataframe

```{r}
# Write code here to read the RDS file
maryland_ppp <- read_rds("data/lab_08.rds")
```

## Answer questions

**Q1** What is the Maryland jurisdiction with the lowest median age, according to the ACS 2019 5-year estimates? You will need to find the variable for median age first in your list of ACS variables.
**A1** Prince George's County (37.1)

```{r}
md <- get_acs(geography = "county",
              variables = c(medage = "B01002_001"),
              state = "MD",
              year = 2019)

md <- md %>% 
  arrange(estimate) %>% 
  top_n(20, estimate)
```

**Q2** Which Maryland counties have a median age of at least 50? Write code to isolate just those jurisdictions.
**A2** Worcester and Talbot counties, taking account their MOE

```{r}
medage_least50 <- md %>% 
  filter(estimate > 50)
```

**Q3** Which Maryland counties have the highest percentage of 5-17 year-olds that speak Spanish at home? You'll need one dataframe with county-level data on the number of 5-17 year-olds that speak Spanish at home and a second dataframe with the total population of 5-17 year-olds by county. For each dataframe, rename the `estimate` column to a meaningful name you've given the ACS variable, remove the `variable` columns in each using select and then join them based on GEOID and NAME. Then you can add a column to calculate the percentage of the total 5-17 population that speaks Spanish at home.
**A3** Prince George's (25.6%) and Montgomery (20.3%)

```{r}
# total population by county
md_youth <- get_acs(geography = "county",
              variables = c(total_youth = "B07001_003"),
              state = "MD",
              year = 2019)

md_youth <- md_youth %>% 
  rename(total_youth = "estimate")

# who speak Span
md_spanish <- get_acs(geography = "county",
                      variables = c(speak_spanish = "B16007_004"),
                      state = "MD", 
                      year = 2019)

md_spanish <- md_spanish %>% 
  rename(young_span = "estimate")

# join
md_youth_span <- md_youth %>% 
  inner_join(md_spanish, by=c("GEOID", "NAME")) %>%  
  rename(variable_youth=variable.x, variable_spanish=variable.y) %>% 
  rename(moe_youth=moe.x, moe_spanish=moe.y)

# percentage
md_youth_span %>% 
  mutate(percentage = 100*(young_span/total_youth)) %>% 
  arrange(desc(percentage))
```

**Q4** Given the counties with the largest percentage of 5-17 year-olds that speak Spanish at home, what stands out to you about the data? How would you describe the top 5 counties? How would you describe Garrett County's numbers?
**A4** I would describe the top 2 counties as the largest counties in Maryland (Prince George's and Montgomery). Those two counties are also adjacent to Washington D.C. The counties will draw families of diplomats, who can speak a foreign language such as Spanish. Then Caroline County houses an ICE Detention Facility. This facility may draw families with family members at the facility. Immigration is also on the rise in Caroline County over the past five years. Caroline, Frederick and Talbot counties are a few of the counties with the lowest cost of living. A low cost of living will draw immigrant families. Garrett County, on the other hand, recorded the least percentage of Black population. This number does not surprise me. Garrett County borders West Virginia, which is commonly known as a white state. Unlike the suburban or urban nature of the top 5 counties, Garrett County is more rural. That county is full of mountainous terrain - perfect for nature hikes.

**Q5** Which 5-digit zip code had the most removed PPP applications, and what is the percentage of Black population in that zip code according to the ACS? To answer this question, you will need to create separate dataframes with ACS data on total population by zip code and Black population by zip code. In each dataframe, rename the `estimate` column and remove the `variable` column. You'll need to combine the two dataframes in order to calculate the percentage, and then you can combine it with the PPP data. Remember: the PPP data may not have only 5-digit zip codes!
**A5** 60620 has the most removed PPP applications. That zip code's Black population is 95.7% of the total percentage.

```{r}
# total pop by zip
maryland_ppp <- maryland_ppp %>%
  mutate(zip5 = str_sub(zip, 1, 5))

# removed apps
top_zips <- maryland_ppp %>% 
  group_by(zip5) %>% 
  summarise(count=n()) %>%
  arrange(desc(count))

# total pop by zip
usa_zips <- get_acs(geography = "zcta",
              variables = c(variables = "B02001_001"))

black_zips <- get_acs(geography = "zcta",
                      variables = c(variables = "B02001_003"))

# join
pop_by_zips <- usa_zips %>% left_join(black_zips, by=c('GEOID'))

pop_by_zips <- pop_by_zips %>% 
  rename(total_pop=estimate.x, black_pop=estimate.y) %>% 
  rename(moe_total=moe.x, moe_black=moe.y)

# percentage
pop_by_zips <- pop_by_zips %>% 
  mutate(percentage = ((black_pop/total_pop)*100))

pop_by_zips %>% 
  arrange(desc(percentage))

# join again
pop_by_zips <- maryland_ppp %>% left_join(pop_by_zips, by=c('zip5'='GEOID'))

pop_by_zips %>% 
  group_by(zip5) %>% 
  summarise(count=n()) %>%
  arrange(desc(count))
```

**Q6** How would you describe to an editor the top 10 zip codes with the most removed applications in terms of Black population? Where are they located, and what do they have in common? Are there any zip codes that seem different, and if so why?
**A6** Most of the top 10 zip codes with the most removed applications are located in and around Chicago, Illinois. The majority start with the numbers 606. All of them also have a high Black population. There is one zip code that stands out - 30349. That zip code is located in Atlanta, Georgia. One commonality with that 30349 zip code and the others is the high Black population. Atlanta records an about 91% Black population of their total population. 

**Q7** What other columns from the PPP data could be useful in reporting on what happened to these removed loans?
**A7** The following columns from the PPP data could be useful in reporting: the number of jobs retained, the date approved and business type. The number of jobs retained can reveal whether the supplied money went to the desired reasons. We can also match the number of jobs matched with the business type. For example, a sole proprietorship should use the money to save one job, whereas a corporation could use the money to save as many as 500 jobs. In particular, questions could arise from limited liabilities companies. CREWFORCE, LLC saved 500 jobs, but ELITE TAX AND NOTARY only saved one. Then the date approved could show reporters when companies/individuals applied for PPP loans. The date approved could reveal under which bill they applied, or even how urgent their request is.
