---
title: "lab_05"
author: "Katie Benzan"
date: "2/24/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

To complete this lab, you need to:
* run existing code as directed (look for **Task**).
* modify existing code as directed (look for **Task**).
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings

You'll need to load two packages for this: the tidyverse and janitor.

**Task** load these two packages.

```{r}
# Turn off scientific notation
options(scipen=999)

library(tidyverse)
library(janitor)

```

## Load Data

You'll need to load three data sets for this:

* The West Virginia slice of the PPP loan data (lab_05.rds).
* A "lookup table" that allows you to translate NAICS (industry) numeric codes to industry titles (naics_codes.csv).
* A table of West Virginia population by county (American Community Survey, 2019 5-year averages) (wv_population_county.csv).

All three data sets are in the data folder.  Write code to load the three in the codeblock below.

**Task** Create a codeblock below this task, then read the data sets in in and assign them to appropriate variable names. There's a fourth data set you'll use in this lab, too, of selected loans in ZIP Code 25401. But there's no need to load it now.

```{r}
wv_ppp <- read_rds("data/lab_05.rds")
naics <- read_csv("data/naics_codes.csv")
wv_pop_by_county <- read_csv("data/wv_population_county.csv")
```


## Answer questions

**Q1.** In the data folder, there is a csv called zip_25401_loan_sample.csv.  It contains a sample of loans from West Virginia ZIP Code 25401.

As we read earlier this semester, [multiple loan applications coming from multiple businesses at the same residential street address](https://www.nytimes.com/2021/08/17/business/ppp-fraud-covid.html) might point to fraud. Or it could alert us to companies that used [multiple corporate entities to get more money than envisioned](https://www.propublica.org/article/different-names-same-address-how-big-businesses-got-government-loans-meant-for-small-businesses) by the original law that authorized the program.   

You are going to examine this data to see if we can find a lot of loans coming from the same address.  Here's the problem: the street address field is pretty messy.  The same address appears with minor variations --  "1003 Sushruta Dr" vs "1003 SUSHRUTA DR" -- that will prevent proper grouping.

First, upload the data into Open Refine and standardize/clean the address field - make sure to make a copy of that column to work on and to try all of the options for clustering. If you've done it properly, you should have 65 discrete addresses.

Then export the data from Open Refine, and move it to the proper folder.

Next, load it in the codeblock below, assigning it to an appropriate variable name.

```{r}
zip_25401_loan <- read_csv("data/zip_25401_loan_data.csv")
```


Then answer these questions:
* What is the street address in this data that has the most loans?
* How many loans are there at that street address?
* What are the names of the businesses at that address?

**A1.** 
 126 E Burke St has the most loans in this data set. 6 loans are recorded for that address. 
The names of those businesses are: DREW HOLDINGS LLC; BRIX27, LLC; ABRAHAM ASHTON; HUB CO-OP LLC; RONIN PROPERTIES LLC; and HUB CO-OP, LLC.
```{r}
zip_25401_loan %>% 
  group_by(address) %>% 
  summarise(
    count = n()
  ) %>% 
  arrange(desc(count))

zip_25401_loan %>% 
  filter(str_detect(address, "^126"))
  
```

**Q2.** Do some web research on the businesses that answered question 1.  

Google the street address.  Google the business names and search for their corporate records on [Open Corporates](https://opencorporates.com/). Be sure to find the website of the name of the company that appears twice in the list of businesses and develop an understanding of what it does.

Based on your research, does it seem suspicious that this collection of businesses all got loans using the same address? Why or why not. Use specific facts identified in your research to support your position.

**A2.**
I do not find the fact that all these businesses listed the same address as suggested. This is because according to Google Maps, 126 E Burke St appears to be a commercial building. There are multiple businesses listed at that address, including The Hub. Interestingly enough, Brix27 - a limited liability company - names Abraham Ashton and Drew Holdings LLC as two officers of their business. Hub Co-Op LLC (another limited liability company) also lists Abraham Ashton as an organizer for the company. This overlap may also explain how multiple businesses list the same address. A limited liability company is a business where the owner is protected from personal responsibility for debts/liabilities. This type of business structure combine a corporation with sole proprietorship. In the case of Brix27, the corporation is specialty food stores. And in the case of The Hub, the corporation is real estate. Therefore, the corporation (i.e. Brix27 or the Hub) can apply for a loan, plus the owner (i.e. Abraham Ashton) can apply for a loan.

**Q3.** Start by using the West Virginia slice of the PPP loan data that you loaded at the start of the lab to create a subset of PPP loans in West Virginia's second largest county (which you can find in wv_population_county). And then use that table you created to answer the following questions:

* Which city in that county had the highest number of loans?
* In that city, which industry title had more loans than any other industry title?

```{r}
wv_pop_by_county %>% 
  arrange(desc(population_2019))

wv_ppp_cleaned <- read_csv("data/wv_ppp_loan_data3.csv")

wv_ppp_cleaned %>% 
  filter(str_detect(project_county_name, "^BER")) %>% 
  group_by(city) %>% 
  summarise(
    count = n()
  ) %>% 
  arrange(desc(count))

wv_ppp_cleaned %>% 
  filter(str_detect(city, "^Mart")) %>% 
  group_by(naics_code) %>% 
  summarise(
    count = n()
  ) %>% 
  arrange(desc(count))
```


Requirement: you MUST export a dataframe of PPP loans from R Studio at some point in the process (not necessarily at the beginning!), load it into Open Refine, clean the city column, export it from Open Refine, and reimport into R Studio. To export data from R Studio, you will use the write_csv() function, which works like this if we're using the wv_population_county dataframe:

```{r}
# write_csv accepts 2 arguments: the dataframe and the path to the file you write out
write_csv(wv_pop_by_county, "data/wv_pop_by_county.csv")

write_csv(wv_ppp, "data/lab_05.csv")
```

Guidance: there are a lot of steps you'll need to take to answer this question. You may find it helpful to write out in English what you plan to do step-by-step before you start writing code.   

**A3.**
Martinsburg is the city in that county with the highest number of loans. And full-service restaurants had more loans (50) than any other industry.

**Q4.** What are your two best hypotheses, which you'd need to confirm with reporting and further analysis, that explain why that industry is at the top of the list?

**A4.**
My first hypothesis to explain why full-service restaurants received the most loans is Americans did not want to eat out during the pandemic. While eating, you must take off your mask. Therefore, eating in an in-closed space is a riskier place to catch covid. Most Americans preferred to order takeout rather than dine in. This tendency leads to less profits, and thus, less cash flow to pay their workers. According to Time Magazine, restaurant sales had diminished by $240 billion in 2020 from their expected numbers. About 80,000 restaurants had to close their doors due to the pandemic and high risk of covid. The National Restaurant Association records the restaurant industry, more than any other industry, has suffered the most substantial losses in sales and jobs during the pandemic. Also, at some point the CDC banned restaurants from even opening their doors.

My second hypothesis is due to the supply chain issues restaurants are experiencing food shortages. According to Time Magazine, a National Restaurant Association survey recorded more than 90% of restaurant operators dealt with supply delays and shortages of popular foods and beverages. Food shortages result in a spike in prices. With the price surge, it’s increasingly difficult to retain employees. Plus, employees tend to avoid work due to childcare problems, risk of infection and presence of unemployment checks during the pandemic. 

**Q5.** Start with a dataframe of loans to all businesses in the city and industry that answered question 3. Answer the following questions:
* What is the name of the business that got the highest approved loan amount?
* How much was it for?
* When was it approved?
* How many jobs does the data say were retained?
* Is there a difference between the business' name in the PPP data and the name its customers know it by? If so, what is that name?
* How many locations does this business have?
* Did one of its locations close during the pandemic, either before or after it got the loan?

Hint: you will not find the answers to the last three questions in the data.  You could call them directly to get that information, but I don't want you to do that for this assignment.  Instead, do some web research. I would start by Googling the company name from the data and looking at the page that comes up for the business from at http://apps.sos.wv.gov/. I would use information I found on that page and use info about the company from Google, the [Wayback machine](https://archive.org/web/) (which lets you look at older versions of a company's website), Yelp, and Facebook.

**A5.**
The highest approved loan was $8,912,202 for Greenbriar Hotel Corporation. It was approved on April 29, 2020, and 500 jobs were retained. 

There is one difference between the business' name in the PPP data and the name its customers know it by. In the PPP data, there is a typo. The PPP data says "Greenbriar" with an "a". But its customers know it by "Greenbrier" with an "e." Also, its customers know it as "American's resort," not necessarily a hotel. A resort implies a full experience with spa treatments and shows. A hotel, on the other hand, implies room and board. The Greenbrier has one location at 101 W Main St, White Sulphur Springs, WV. That Greenbrier location closed for two months due to covid-19. On May 22, 2020, it reopened. Thus, It closed around March 20, meaning the resort closed before their loan was approved in April.
```{r}
wv_ppp_cleaned %>% 
  summarise(
    max(current_approval_amount)
  )

wv_ppp_cleaned %>% 
  filter(str_detect(current_approval_amount, "8912202")) 
```

