---
title: "lab_07"
author: "katie benzan"
date: "3/10/22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## About this lab

read_html - given a url, reads the html
html_table - given html, finds the <table> objects
html_elements - given html, you supply one or more tags OR an xpath expression
  ex. my_stuff <- url %>% 
    read_html() %>% 
    html_elements(`p a`) #this finds all the <p> tags
html_text - the contents of any tag

1. read the html
2. find the elements i need
3. grad the dataframe and put it in a container of all of my data

To complete this lab, you need to:
* write code in empty codeblocks provided to answer questions included (look for **Q**).
* write out the answer in the form of a complete sentence in the space given (look for **A**).

When you are finished, commit changes and push to your personal GitHub repo, then submit the URL to this document on ELMS.

## Load libraries and establish settings
**Task**: Load rvest, janitor and the tidyverse
```{r}
# Turn off scientific notation
options(scipen=999)

# Load the tidyverse,rvest,janitor
library(rvest)
library(tidyverse)
library(janitor)
```


Q1. How many individual cases has the U.S. Department of Justice filed against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  An example of one case that appears on this page is "U.S. v. Richard Ayvazyan et al". To answer this question, you will need to use rvest to scrape the data on this single webpage into a dataframe that lists only case names. Hint: you will need to make use of html_elements() and html_text() -- a function that extracts text inside of an html tag -- for this.
A1. 100 cases

```{r}
# store URL
fraud_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

# read html
fraud_cases <- fraud_url %>% 
  read_html()

# find elements
fraud_cases <- fraud_url %>% 
  read_html() %>% 
  html_elements('li b') %>% 
  html_text('li b')

# dataframe
fraud_cases <- as.data.frame(fraud_cases)

# clean
fraud_cases <- fraud_cases %>% 
  clean_names() %>% 
  distinct()

```

Q2. In how many individual judicial districts has the U.S. Department of Justice filed cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud?  Note: an example of a judicial district is "Southern District of Florida". You will need to use rvest scrape the data on this single webpage into a dataframe.
A2. 31 districts.

```{r}
# store URL
fraud_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

# read html
fraud_districts <- fraud_url %>% 
  read_html()

# find elements
fraud_districts <- fraud_url %>% 
  read_html() %>% 
  html_elements('b i') %>% 
  html_text('b i')

# dataframe
fraud_districts <- as.data.frame(fraud_districts)

# clean
fraud_districts <- fraud_districts %>% 
  clean_names() %>% 
  distinct()
```

Q4. The website (https://www.justice.gov/criminal-fraud/cares-act-fraud) shows that the DOJ filed more cases in the Southern District of Florida than in any other district. One of those cases was filed against someone named "Diamond Blue Smith". Who is Smith, and what was he accused of, according to the criminal complaint? If you were an editor, would you have assigned a reporter to write a story about this case when the complaint was filed in court? Why or why not?
A4. Smith is accused of wire fraud, bank fraud and conspiracy/attempt to commit wire fraud and bank fraud. Smith obtained two PPP loans for his companies - Throwbackjerseys.com, LLC ("Throwbackjerseys.com"), and BlueStarRecords, LLC ("BlueStarRecords"). Smith also teamed up with other applicants to submit at least 90 fake applications that totaled more than $24 million. Almost half of those loans were approved, scamming the federal government out of about $17.4 million. Due to these figures, I would assign a reporter to this case. Smith was responsible and the mind behind an enormous deception of the federal government. 

Q5. In what percentage of all judicial districts has the U.S. Department of Justice filed cases against people accused of fraud related to the PPP or EIDL loan program, according to the DOJ website listing those cases: https://www.justice.gov/criminal-fraud/cares-act-fraud? In answering this question, you should also produce a list of judicial districts where DOJ has NOT filed a case, according to this site.  Note: to answer this question, you will need to scrape a table of all district courts on this up-to-date Wikipedia page under the heading "Active Courts": https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts  
A5. In 33% of all judicial districts, the DOJ filed PPP fraud cases. The DOJ did not file a case in 67 districts. 


```{r}
# store URL
fraud_url <- "https://www.justice.gov/criminal-fraud/cares-act-fraud"

# read html
fraud_districts <- fraud_url %>% 
  read_html()

# find elements
fraud_districts <- fraud_url %>% 
  read_html() %>% 
  html_elements('b i') %>% 
  html_text('b i')

# dataframe
fraud_districts <- as.data.frame(fraud_districts)

# clean
fraud_districts <- fraud_districts %>% 
  clean_names() %>% 
  distinct()

# all judicial courts
allcourts_url <- "https://en.wikipedia.org/wiki/List_of_United_States_district_and_territorial_courts"

# read html
district_courts <- allcourts_url %>% 
  read_html()

# find elements
district_courts <- allcourts_url %>% 
  read_html() %>% 
  html_table()

# dataframe
district_courts <- district_courts[[3]] %>% 
  clean_names() %>% 
  distinct()

# new column
fraud_districts <- fraud_districts %>% mutate(doj_filing = "filed")
fraud_districts <- fraud_districts %>% 
  rename(region = fraud_districts)

# combine dataframes
doj_courts_noppp <- district_courts %>% 
  anti_join(district_courts, fraud_districts, by = c("region" = "fraud_districts)) 
  
# percentage = 100 * (31/94)
```
Q6. What might explain why, according to this site, the DOJ has filed PPP-related fraud cases in less than half of the country's judicial districts?
A6. 

Q7. Which state had the most approved PPP loans per 100,000 population? [This web page](https://dwillis.github.io/jour472files/ppp_scraping_example/index.html) has links to 52 individual web pages, one for each state (plus Washington, D.C. and Puerto Rico). Each of those web pages contains a one-row html table that has the name of the state, the number of approved loans, and the 2019 population. Here's an example for [Alabama](https://dwillis.github.io/jour472files/ppp_scraping_example/states/alabama.html). You'll need to loop over the individual state urls, scrape each individual page and combine the information on each page into a single dataframe to answer this question. Don't forget to calculation the loans per 100,000.
A7.
```{r}
# url
states_url <- "https://dwillis.github.io/jour472files/ppp_scraping_example/index.html"

# read html
states <- states_url %>% 
  read_html() %>% 
  
state_list <- states_url %>% 
  read_html() %>% 
  html_table

state_list <- state_list[[1]] %>% 
  clean_names()

# number of rows
for(row_number in 1:nrow(state_list)) {

    # Keep only the row for a given row number, get rid of every other row
    each_row_df <- state_list %>%
      slice(row_number)
    
     # Define url of page to get
    url <- each_row_df$url
    
    # Define id of table to ingest
    xpath_state_table <- paste0('//*[@id="',each_row_df$sector_xpath_id,'"]')
}
```
